A comparison of design-based and model-based approaches for spatial data.

### 2021-08-09 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
coverage and pros/cons updates

combo_data <- purrr::map_df(files,
                            ~read_csv(.x) %>% mutate(filename = .x)) %>%
  separate(filename, into = c("junk", "error_dep", "dist", "n", "location"),
           sep = "_") %>%
  select(-junk)
combo_data <- combo_data %>% mutate(sim = interaction(error_dep, dist, n, location))

@@

4. Take-home messages

@@

__Base Simulations__

* both good: correctly specified model with high correlation (we did this in Table \ref{tab:sim_one})
* break model: highly non-normal errors with small sample size
* break design: small area estimation

__Simulation Discussion Questions__

* model-based: how should sample be drawn? should locations be fixed?
* change n or sampling fraction?

__Other Base Settings?__

* both good?: misspecified covariance model with high correlation
* break both? non-gaussian areas with smaller sample size

@@

>>>>

# mike chunk
# combo_data <- purrr::map_df(files,
#                             ~read_csv(.x) %>% mutate(filename = .x)) %>%
#   separate(filename, into = c("junk", "junk2", "error_dep", "dist", "n", "location"),
#            sep = "_") %>%
#   select(-junk, -junk2)

# matt chunk
combo_data <- purrr::map_df(files,
                            ~read_csv(.x) %>% mutate(filename = .x)) %>%
  separate(filename, into = c("junk", "error_dep", "dist", "n", "location"),
           sep = "_") %>%
  select(-junk)

combo_data <- combo_data %>% mutate(sim = interaction(error_dep, dist, n, location))

@@

4. Coverage

For Gaussian errors, coverage for all approaches tended to be near 0.95. There was less between-approach deviation in coverages for random locations compared to grid locations. Generally, the larger the skew, the worse the coverage,and the larger the sample size, the better the coverage. Design GRTS (local neighborhood variance) tended to slightly undercover, a result Tony was familiar with.

```{r}
ggplot(data = combo_data, aes(x = sim, y = coverage, colour = approach)) +
  geom_hline(yintercept = 0.95) +
  geom_point() +
  coord_flip()
```

5. Take-home messages

@@

@@

* Pros of Design-Based (items we are not exploring): computationally efficient, few assumptions, more naturally handles binary data,

* Pros of Model-Based (items we are not exploring): covariate inference, more efficient small-area estimation, model selection?

### 2021-08-09 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
knitted writeup
coverage and pros/cons updates

% Pandoc citation processing
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

@@

Though the design-based and model-based approaches apply to statistical
inference in a broad sense, we focus on comparing these approaches for
spatial data. We define spatial data as variables measured at specific
geographic locations. De Gruijter and Ter Braak (1990) give an early
comparison of design-based and model-based approaches for spatial data,
quashing the belief that design-based approaches could not be used for
spatially correlated data. Thereafter, several comparisons between
design-based and model-based for spatial data have been considered, but
they tend to compare design-based approaches that ignore spatial
locations to model-based approaches (Brus and De Gruijter, 1997; Ver
Hoef, 2002; Ver Hoef, 2008). Cooper (2006) review the two approaches in
an ecological context before introducing a ``model-assisted'' variance
estimator that combines aspects from each approach. In addition to
Cooper (2006), there has been substantial research and development into
estimators that use both design and model-based principles (see e.g.
Cicchitelli and Montanari (2012), Chan-Golston et al. (2020) for a
Bayesian approach, and Sterba (2009)). More recent overviews include
Brus (2020) and Wang et al. (2012), but no numerical comparison has been
made between design-based approaches that incorporate spatial locations
and model-based approaches.

@@

\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-4-1.pdf}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{3}
\tightlist
\item
  Take-home messages
\end{enumerate}

@@

\textbf{Base Simulations}

\begin{itemize}
\tightlist
\item
  both good: correctly specified model with high correlation (we did
  this in Table \ref{tab:sim_one})
\item
  break model: highly non-normal errors with small sample size
\item
  break design: small area estimation
\end{itemize}

\textbf{Simulation Discussion Questions}

\begin{itemize}
\tightlist
\item
  model-based: how should sample be drawn? should locations be fixed?
\item
  change n or sampling fraction?
\end{itemize}

\textbf{Other Base Settings?}

\begin{itemize}
\tightlist
\item
  both good?: misspecified covariance model with high correlation
\item
  break both? non-gaussian areas with smaller sample size
\end{itemize}

@@

\begin{longtable}[]{@{}lrrrrr@{}}
\caption{\label{tab:appliedtab} Table XXX. Application of design-based
and model-based approaches to the NLA data set on mercury
concentration.}\tabularnewline
\toprule
Approach & Realized Mean & Estimate & SE & 95\% LB & 95\% UB \\
\midrule
\endfirsthead
\toprule
Approach & Realized Mean & Estimate & SE & 95\% LB & 95\% UB \\
\midrule
\endhead
Design IRS & 103.2 & 112.7 & 8.8 & 95.4 & 129.9 \\
Model IRS & 103.2 & 110.5 & 7.9 & 95.0 & 125.9 \\
Design GRTS & 103.2 & 101.8 & 6.1 & 89.8 & 113.7 \\
Model GRTS & 103.2 & 102.3 & 5.9 & 90.8 & 113.9 \\
\bottomrule
\end{longtable}

@@

>>>>

% Pandoc citation processing

@@

Though the design-based and model-based approaches apply to statistical
inference in a broad sense, we focus on comparing these approaches for
spatial data. We define spatial data as variables measured at specific
geographic locations. De Gruijter and Ter Braak (1990) give an early
comparison of design-based and model-based approaches for spatial data,
quashing the belief that design-based approaches could not be used for
spatially correlated data. Thereafter, several comparisons between
design-based and model-based for spatial data have been considered, but
they tend to compare design-based approaches that ignore spatial
locations to model-based approaches (Brus and De Gruijter, 1997; Ver
Hoef, 2002, 2008). Cooper (2006) review the two approaches in an
ecological context before introducing a ``model-assisted'' variance
estimator that combines aspects from each approach. In addition to
Cooper (2006), there has been substantial research and development into
estimators that use both design and model-based principles (see
e.g.~Cicchitelli and Montanari (2012), Chan-Golston et al. (2020) for a
Bayesian approach, and Sterba (2009)). More recent overviews include
Brus (2020) and Wang et al. (2012), but no numerical comparison has been
made between design-based approaches that incorporate spatial locations
and model-based approaches.

@@

\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-4-1.pdf}
\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-4-2.pdf}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{3}
\tightlist
\item
  Coverage
\end{enumerate}

For Gaussian errors, coverage for all approaches tended to be near 0.95.
There was less between-approach deviation in coverages for random
locations compared to grid locations. Generally, the larger the skew,
the worse the coverage,and the larger the sample size, the better the
coverage. Design GRTS (local neighborhood variance) tended to slightly
undercover, a result Tony was familiar with.

\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-6-1.pdf}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{4}
\tightlist
\item
  Take-home messages
\end{enumerate}

@@

@@

\begin{longtable}[]{@{}lrrrrr@{}}
\caption{\label{tab:appliedtab} Table XXX. Application of design-based
and model-based approaches to the NLA data set on mercury
concentration.}\tabularnewline
\toprule
Approach & Realized Mean & Estimate & SE & 95\% LB & 95\%
UB\tabularnewline
\midrule
\endfirsthead
\toprule
Approach & Realized Mean & Estimate & SE & 95\% LB & 95\%
UB\tabularnewline
\midrule
\endhead
Design IRS & 103.2 & 112.7 & 8.8 & 95.4 & 129.9\tabularnewline
Model IRS & 103.2 & 110.5 & 7.9 & 95.0 & 125.9\tabularnewline
Design GRTS & 103.2 & 101.8 & 6.1 & 89.8 & 113.7\tabularnewline
Model GRTS & 103.2 & 102.3 & 5.9 & 90.8 & 113.9\tabularnewline
\bottomrule
\end{longtable}

@@

\begin{itemize}
\item
  Pros of Design-Based (items we are not exploring): computationally
  efficient, few assumptions, more naturally handles binary data,
\item
  Pros of Model-Based (items we are not exploring): covariate inference,
  more efficient small-area estimation, model selection?
\end{itemize}
